```{r}
library(tidyverse)
library(nycflights13)
```
#Produce a scatter plot of displ vs hwy.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

#Add another aesthetic: map a continuous variable to colour.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = cty))
```

#Map the same continuous variable to size.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = cty, size = cty))
```

#Map the same continuous variable to shape.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(fill = "red")
```

```{r}
head(flights)
```

```{r}
select(flights, c("dep_time", "dep_delay", "arr_time", "arr_delay"))
```
```{r}
select(flights, c("dep_time", "dep_delay", "arr_time", "arr_delay", "arr_delay"))
```
```{r}
str(flights)
```
#Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights1 = mutate(flights,
       dep_time = ((dep_time %/% 100) * 60 + (dep_time %% 100)) %% 1440,
       sched_dep_time = ((sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100)) %% 1440)
```

```{r}
head(flights)
```


```{r}
flights %>% slice(c(1:5, 8, 15:20))
```
#Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}

```

#Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
```{r}
flightsdep_delay <- mutate(flightsnew, dep_time_min = tim2min(dep_time), sched_dep_time_min = tim2min(sched_dep_time), act_dep_delay = dep_delay-dep-time + sched_dep_time_min)
fliter(flightsdep_delay, act_dep_delay != 0)
```



